\documentclass[12pt]{article}

% Import all needed packages
\usepackage[letterpaper, margin=1in]{geometry}
\usepackage{setspace}

% Set double spacing
\doublespacing

% Bibliography Setup
\usepackage[backend=bibtex]{biblatex}
\addbibresource{bibliography.bib}

% Setup cover page.
\title{Applications of Machine Learning in Enzyme Design}
\author{Jonah Nichols}
\date{\today}

\begin{document}
\maketitle
\section{Introduction}
Plastic is an incredibly versatile and cheap material. It is used in nearly every industry,
from packaging to consumable goods to medicine to research \cite{geyer2017production}.
Estimates for global plastic production range from 335 MT to 380 MT \cite{drzyzga2019plastic,
geyer2017production,leal2019overview}. This quantity is high due to its high versatility
and durability. After its usage though, plastic lingers in range from a few years
to decades. Polyethylene terephthalate, for example, takes one year to degrade one
$\mu$m of plastic from the outside in \cite{chamas2020degradation}

% Maybe a paragraph on how long plastic takes to degrade in a natural environment?
% Maybe include that in this paragraph at the end?
% Maybe quite focusing on the labor-intensive quantification?
Plastic waste management can be done in a few different ways, including but not
limited to placing it into a landfill or burning it.
The more labor intensive solution is to burn the plastic to help regain some of
the energy lost from its production. This comes with the harmful fumes which can
wreak havoc amongst various organs and body systems \cite{agnes2016environmental}.
Placing it into the ocean and landfill are both quick and easy, yet have vast ecological
impacts for marine creatures living in those environments.

% Maye discuss the impacts stated above?

Bioremediation is a more viable solution for plastic waste management as it converts
harmful polutants into harmless byrpoducts. Bioremediation
can be helpful because it is cheaper and can provide a more long-term low-tech solution \cite{vidali2001bioremediation}.
While it is true that bioremediation is slower and incredibly plastic specific,
%maybe another citation here?
it's important to remember that enzyme design could help improve upon the specificity
by designing enzymes to degrade a variety polymers. Not only could this bioremediate
plastic-contaminated samples, it could also handle other chemical compounds \cite{karigar2011role}.

Enzyme design could be expanded even further though. \textit{Most} chemical reactions could
be catalyzed by enzymes. The issue with the current enzyme design process is that it is rather
lengthy and difficult to use \cite{richter2011novo}. The goal of this research project was to
use a machine learning method on ester hydrolysis reactions to create enzymes that catalyze them.
This is in an attempt to create an easier pipeline for creating enzymes.

\section{Methods}
\subsection{Data Acquisition}
RheaDB was utilized to find ester hydrolysis reactions according to the Enzyme Classification
(EC) standards, specifically any enzyme with the classification 3.1.x.x. From selected reactions, Rhea's cross references were utilized
to obtain sequences for the various chemical reactions from UniprotKB. For every
sequence-reaction pair, the Morgan Fingerprint Difference of the reaction was created using
RDKit's built-in fingerprinting function. Due to a lack of known directional data,
unknown directional enzymes were assumed to catalyse the left-to-right version of
the reaction. Next the fingerprint was converted into a PyTorch tensor for import
into the machine learning algorithm.

For every sequence, a one-hot encoding of the sequence was built to create a suitable
output for the reaction. Each amino acid was represented by an integer.
These sequences were then converted into PyTorch tensors
for input into the machine learning algorithm paired up with its corresponding reaction fingerprint.
These pairs were split randomly into a ratio of 75\% training and 25\% test datasets

\subsection{Neural Network Design}
Given inexperience in the subject, the neural network was designed to a very low
degree of confidence. The input layer was the reaction fingerprint previously generated.
Three repetitions of linear layers followed by ReLU activation layers expanded the
fingerprint out to 7192 bits, corresponding to the longest sequence found. A final
linear layer presented the output of the model.

\subsection{Training and Analysis}
For training, the inputs and outputs were batched to 64 samples and shuffled to
limit overfitting. The L1 Loss function was used to determine the error of the
model. Stochastic gradient descent was utilized to optimize the model.
The training sequence was run for 100 epochs to get an optimized model.

Due to time constraints, the model's learning capabilities and pattern recognition
were not determined.

\section{Results}
Of the 9890 sequences queried from UniprotKB with the enzyme class of 3.1.x.x, 19084
reaction-sequence pairs were generated and used for training in the machine learning
model. 

In a visual examination of the output the neural network gave, it became apparent that
the neural network had not learned much of the data. The output given was to be a
one-hot encoded sequence in number form, to be converted back to a proper sequence. Unfortunately
the output received was a sequence of numbers between 12-14 and then a gradual descent
to 7's, presumably the "pad" character created to sync 

\section{Discussion}
Future work on the subject is key to enhancing our ability to create and use enzymes
to catalyze reactions. The most arbitrary extension of this work is in rewriting
the featurization of the enzymes to more resemble other machine learning projects.
Rather than represent each amino acid as a number, represent it as a 21-bit tensor.
Each bit would represent a different amino acid being turned on. 

Another more laborious route for the project is to move towards a 3D-based approach
opposed to the current sequence-based approach. This route is more difficult due
to the conversion from a structure to a one-dimensional input for a neural network.
Before the fasta-approach branch it was intended as a structure-based approach but
processing times were too long for a limited time scale.

Another another pathway to take would be returning the project to a more typical
enzyme design process. This would involve completely refactoring all of the work
done so far, but may offer better end results. The process, described briefly, is
beginning with the simulated transition state using density-functional theory (DFT)
and building up the enzyme around said transition state.

Regardless, continuing work on enzyme design is irrefutably important. Creating
enzymes from scratch has possibilities ranging from new antibiotics to ecological
benefits to commercial benefits. The possibilities are endless.

\printbibliography

\end{document}